Customizing Software by Direct Manipulation of Structured Data

### 2020-05-21 paper.tex
more edits from dnj

@@

It is reasonable to choose imperative programming as the model for
building customizations. Turing-Complete programming provides a high
ceiling for possible customizations, and the a sequence of commands is a
natural fit for automations which simulate a series of steps taken by
the user.

@@

We have used Wildcard to build customizations for many real websites. So
far, most usage has come from members of the project team, although we
have had a few external users try the system. Here we offer our
reflections on using the system, focused on two key questions:

@@

\emph{todo}

@@

In order for our idea of third-party customization through Wildcard to
succeed, it is important that creating usable adapters for existing
websites takes minimal effort.

To test the viability of writing DOM scraping adapters, we built
adapters for 10 websites, including transactional sites like Amazon and
Uber Eats, media consumption sites like Hacker News and Youtube, and a
medical results entry platform managed by a member of our research
group. The adapters ranged from 36 to 117 lines of code, averaging 68
lines. In addition, an external developer unaffiliated with the project
contributed one adapter, designed to sort the Github page listing a
user's repositories, and described the experience as ``very
straightforward.''

>>>>

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

@@

Imperative programming is a reasonable choice as the model for building
customizations. Turing-complete programming provides a high ceiling for
possible customizations, and a sequence of commands is a natural fit for
automations which simulate a series of steps taken by the user.

@@

We used Wildcard to customize 11 websites, including transactional sites
like Amazon and Uber Eats, media consumption sites like Hacker News and
Youtube, and a medical results entry platform managed by a member of our
research group. (\emph{todo: summarize all of these in a table}) So far,
most usage has come from members of the project team, although we have
had a few external users try the system.

Here we offer our reflections on using the system, focused on two key
questions:

@@

@@

In order for third-party customization through Wildcard to succeed, it
is important that creating usable adapters for existing websites takes
minimal effort.

Most of our adapters were created by members of our team. However, an
external developer unaffiliated with the project contributed one
adapter, designed to sort the Github page listing a user's repositories.
They described the experience as ``very straightforward.''

\begin{Shaded}
\begin{Highlighting}[]
\StringTok{\textquotesingle{}use strict\textquotesingle{}}\OperatorTok{;}
\ImportTok{import} \OperatorTok{\{}\NormalTok{ urlMatches}\OperatorTok{,}\NormalTok{ extractNumber }\OperatorTok{\}} \ImportTok{from} \StringTok{"../utils"}\OperatorTok{;}
\ImportTok{import} \OperatorTok{\{}\NormalTok{ createDomScrapingAdapter }\OperatorTok{\}} \ImportTok{from} \StringTok{"./domScrapingBase"}

\KeywordTok{const}\NormalTok{ GithubAdapter }\OperatorTok{=} \AttributeTok{createDomScrapingAdapter}\NormalTok{(}\OperatorTok{\{}
  \DataTypeTok{name}\OperatorTok{:} \StringTok{"Github"}\OperatorTok{,}
  \DataTypeTok{enabled}\OperatorTok{:}\NormalTok{ () }\KeywordTok{=>} \AttributeTok{urlMatches}\NormalTok{(}\SpecialStringTok{/github.com.}\SpecialCharTok{*}\SpecialStringTok{tab=repositories/}\NormalTok{)}\OperatorTok{,}
  \DataTypeTok{attributes}\OperatorTok{:}\NormalTok{ [}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"name"}   \OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{\},}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"stars"}  \OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"numeric"}\OperatorTok{\},}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"forks"}  \OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{\},}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"updated"}\OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{\},}
    \CommentTok{// }\AlertTok{TODO}\CommentTok{ datetime type would be nice? not everything has ISO formatted strings}
\NormalTok{  ]}\OperatorTok{,}
  \AttributeTok{scrapePage}\NormalTok{() }\OperatorTok{\{}
    \ControlFlowTok{return} \VariableTok{Array}\NormalTok{.}\AttributeTok{from}\NormalTok{(}\VariableTok{document}\NormalTok{.}\AttributeTok{querySelectorAll}\NormalTok{(}\StringTok{"li.source"}\NormalTok{)).}\AttributeTok{map}\NormalTok{(el }\KeywordTok{=>} \OperatorTok{\{}
      \KeywordTok{let}\NormalTok{ name\_el }\OperatorTok{=} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{\textquotesingle{}a[itemprop="name codeRepository"]\textquotesingle{}}\NormalTok{)}
      \KeywordTok{let}\NormalTok{ name }\OperatorTok{=} \VariableTok{name\_el}\NormalTok{.}\VariableTok{textContent}\NormalTok{.}\AttributeTok{trim}\NormalTok{()}

      \KeywordTok{let}\NormalTok{ stars\_el }\OperatorTok{=} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{\textquotesingle{}*[href*="/stargazers"\textquotesingle{}}\NormalTok{)}
      \KeywordTok{let}\NormalTok{ stars }\OperatorTok{=} \AttributeTok{extractNumber}\NormalTok{(stars\_el}\OperatorTok{,} \DecValTok{0}\NormalTok{)}

      \KeywordTok{let}\NormalTok{ forks\_el }\OperatorTok{=} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{\textquotesingle{}*[href*="/network/members"]\textquotesingle{}}\NormalTok{)}
      \KeywordTok{let}\NormalTok{ forks }\OperatorTok{=} \AttributeTok{extractNumber}\NormalTok{(forks\_el}\OperatorTok{,} \DecValTok{0}\NormalTok{)}

      \KeywordTok{let}\NormalTok{ lang\_el }\OperatorTok{=} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{\textquotesingle{}*[itemprop="programmingLanguage"]\textquotesingle{}}\NormalTok{)}
      \CommentTok{// some repos don\textquotesingle{}t have language set}
      \KeywordTok{let}\NormalTok{ lang }\OperatorTok{=}\NormalTok{ lang\_el }\OperatorTok{==} \KeywordTok{null} \OperatorTok{?} \KeywordTok{null}\NormalTok{ : }\VariableTok{lang\_el}\NormalTok{.}\VariableTok{textContent}\NormalTok{.}\AttributeTok{trim}\NormalTok{()}

      \KeywordTok{let}\NormalTok{ updated\_el }\OperatorTok{=} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{\textquotesingle{}relative{-}time\textquotesingle{}}\NormalTok{)}
      \KeywordTok{let}\NormalTok{ updated }\OperatorTok{=} \VariableTok{updated\_el}\NormalTok{.}\AttributeTok{getAttribute}\NormalTok{(}\StringTok{\textquotesingle{}datetime\textquotesingle{}}\NormalTok{)}

      \ControlFlowTok{return} \OperatorTok{\{}
          \DataTypeTok{id}\OperatorTok{:}\NormalTok{ name}\OperatorTok{,}
          \DataTypeTok{rowElements}\OperatorTok{:}\NormalTok{ [el]}\OperatorTok{,}
          \DataTypeTok{dataValues}\OperatorTok{:} \OperatorTok{\{}
              \DataTypeTok{name}\OperatorTok{:}\NormalTok{ name}\OperatorTok{,}
              \DataTypeTok{stars}\OperatorTok{:}\NormalTok{ stars}\OperatorTok{,}
              \DataTypeTok{forks}\OperatorTok{:}\NormalTok{ forks}\OperatorTok{,}
              \DataTypeTok{updated}\OperatorTok{:}\NormalTok{ updated}\OperatorTok{,}
          \OperatorTok{\},}
      \OperatorTok{\}}
    \OperatorTok{\}}\NormalTok{)}
  \OperatorTok{\},}
  \DataTypeTok{onRowSelected}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \OperatorTok{\{}
      \VariableTok{row}\NormalTok{.}\VariableTok{rowElements}\NormalTok{.}\AttributeTok{forEach}\NormalTok{(el }\KeywordTok{=>} \OperatorTok{\{}
          \ControlFlowTok{if}\NormalTok{ (}\VariableTok{el}\NormalTok{.}\AttributeTok{style}\NormalTok{) }\OperatorTok{\{}
              \VariableTok{el}\NormalTok{.}\AttributeTok{style}\NormalTok{[}\StringTok{"background{-}color"}\NormalTok{] }\OperatorTok{=} \StringTok{"\#def3ff"}
          \OperatorTok{\}}
      \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
      \VariableTok{row}\NormalTok{.}\AttributeTok{rowElements}\NormalTok{[}\DecValTok{0}\NormalTok{].}\AttributeTok{scrollIntoView}\NormalTok{(}\OperatorTok{\{} \DataTypeTok{behavior}\OperatorTok{:} \StringTok{"smooth"}\OperatorTok{,} \DataTypeTok{block}\OperatorTok{:} \StringTok{"center"} \OperatorTok{\}}\NormalTok{)}
  \OperatorTok{\},}
  \DataTypeTok{onRowUnselected}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \OperatorTok{\{}
      \VariableTok{row}\NormalTok{.}\VariableTok{rowElements}\NormalTok{.}\AttributeTok{forEach}\NormalTok{(el }\KeywordTok{=>} \OperatorTok{\{}
          \ControlFlowTok{if}\NormalTok{(}\VariableTok{el}\NormalTok{.}\AttributeTok{style}\NormalTok{) }\OperatorTok{\{}
              \VariableTok{el}\NormalTok{.}\AttributeTok{style}\NormalTok{[}\StringTok{"background{-}color"}\NormalTok{] }\OperatorTok{=} \VerbatimStringTok{\textasciigrave{}\textasciigrave{}}
          \OperatorTok{\}}
      \OperatorTok{\}}\NormalTok{)}
  \OperatorTok{\},}
\OperatorTok{\}}\NormalTok{)}

\ImportTok{export} \ImportTok{default}\NormalTok{ GithubAdapter}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

The adapters for our test sites ranged from 36 to 117 lines of code,
averaging 68 lines.
